---
title: "NHL Shots per Goal Model"
author: "Sasank Vishnubhatla"
date: "January 4, 2019"
runtime: shiny
output: html_document
---

```{r, echo = FALSE}
rm(list = ls())

seed = 15224
set.seed(seed)
```

In the NHL, goalies are arguably one of the most teammembers. Most deep postseason runs rely on consistent goalie play. However, being able to compare goalies tends to be difficult, since consistency is extremely hard for a goalie. Few goalies have been consistent their entire career. Of the modern day goalies, very few exemplify extreme consistency: Henrik Lundqvist, Marc-Andre Fleury, Roberto Luongo and Jonathan Quick to name a few.

In this document, I'll be detailing an analysis and seeing how it pertains to consistency. The data used has been taken from [MoneyPuck](moneypuck.com).

# Data Formatting

So, let's load in our data:

```{r}
data_2017 = read.csv("data/2017.csv")
```

We'll be starting with the 2017-2018 season. Since we have an extremely large amount of data, we need to clean it up. We'll only be looking at regular season games, initially, so we can start by subsetting our data on that.

```{r}
regular_season = subset(data_2017, isPlayoffGame == 0)
```

Some important columns we need to keep are:

* xCord
* yCord
* xCordAdjusted
* yCordAdjusted
* shotAngle
* shotAngleAdjusted
* shotDistance
* playerPositionThatDidEvent
* goalieIdForShot
* goalieNameForShot
* shooterPlayerId
* shooterName

We'll rename some of the columns, so here is a handy table of the old column names and the new column names.

| Old Column Name | New Column Name |
| --------------- | --------------- |
| xCord | x |
| yCord | y |
| xCordAdjusted | x_adj |
| yCordAdjusted | y_adj |
| shotAngle | angle |
| shotAngleAdjusted | angle_adj |
| goal | goal |
| goalieIdForShot | goalie_id |
| goalieNameForShot | goalie_name |
| shooterPlayerId | skater_id |
| shooterName | skater_name |
| playerPositionThatDidEvent | pos |

Now, let's create a new dataframe with just those columns.

```{r}
analysis = data.frame(x = regular_season$xCord,
					  y = regular_season$yCord,
					  x_adj = regular_season$xCordAdjusted,
					  y_adj = regular_season$yCordAdjusted,
					  angle = regular_season$shotAngle,
					  angle_adj = regular_season$shotAngleAdjusted,
					  goal = regular_season$goal,
					  goalie_id = regular_season$goalieIdForShot,
					  goalie_name = regular_season$goalieNameForShot,
					  skater_id = regular_season$shooterPlayerId,
					  skater_name = regular_season$shooterName,
					  pos = regular_season$playerPositionThatDidEvent)
```

With this `analysis` dataframe, we can start looking at a proposed statistic: shots per goal.

# Shots Per Goal

Shots per goal is a new statistic. Initially, when I had this idea, it was only for goalies. [This article](http://sasankvishnubhatla.net/nhl-goalie-shot-analysis/) details an initial attempt at analyzing what a shot per goal means. In that initial analysis, I looked at several goalies and tried to determine if there was any significance in what a shot per goal meant. However, now, with further thought, we'll be looking at what a shot per goal means to more than just goalies.

## Goalie Aspect

Let's start be looking at a few select goalies from the 2017-2018 season:

1. Braden Holtby
2. Marc-Andre Fleury
3. Connor Hellebuyck
4. Carey Price
5. Andrei Vasilevskiy
6. Robin Lehner
7. Antti Niemi
8. Jaroslav Halak
9. Peter Budaj
10. Jacob Markstrom

To start, we need a function that will take a goalie's name and return all their data. So, let's write our specified subset function:

```{r}
getGoalieData <- function(data, name) {
	subset(data, goalie_name == name)
}
```

So, let's now start with Brayden Holtby:

```{r}
holtby = getGoalieData(analysis, "Braden Holtby")
```

## Skater Aspect

## Positional Aspect

## Locational Aspect

# Visualizations